Julia for HPC: In Situ Data Analysis with Julia for Climate Simulations at Large Scale

### 2024-06-18 paper/paper.tex
Update paper.tex

The Julia interface mainly defines the format of input data, so all in situ Julia data analysis methods can be plugged-in without additional E3SM compilation, thanks to Julia's Just-in-time (JIT) compilation.
As discussed previously, another key design of the Julia interface is to correctly convert the Fortran MPI communicator to Julia MPI communicator, which enables the algorithm developers to write their analysis routines using the same EAM MPI communicator. Due to the low-level MPI implementations, different MPI libraries (i.e., OpenMPI, MPICH, and MVAPICH) require specific converters; the following code snippet shows the converter implementations~\cite{gabriel2004open,panda2013mvapich,gropp1996user}.

@@

\subsection{PCA Results}
Over the one-year simulation run, we extracted the top three spatial principal components of the surface temperature variable using the TributaryPCA method; the principal components were updated in a streaming fashion at each simulation time step (6-hour time steps).
Figure \ref{fig:pca_results} shows spatial plots of these three principal components. 
Each component indicates a different spatial pattern of warming (red color) and cooling (blue color) extracted from the simulation data.
In particular, the first component indicates a temperature differential between equatorial and polar regions; because the mean was not removed, this component can be interpreted as the mean pattern across the time steps. 
The second component indicates a temperature differential between much of the Eastern hemisphere and the Americas and parts of Africa. 
The third component indicates a temperature differential between Africa and Europe and other regions.
While these components were extracted from a fairly short simulation run, the results demonstrate the ability of TributaryPCA to extract important climate patterns while running in a distributed in situ environment, without saving any raw data for post-processing.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/insitu_pca.png}
    \caption{First three spatial principal components obtained by in situ TributaryPCA method. Within each component, the white color indicates no change to surface temperature, while blue and red indicate temperatures deviating in different directions from zero. The first principal component appears to indicate the overall difference between the poles and the equator. The second principal component indicates differences between the Eastern hemisphere and Africa/the Americas, while the third component indicates differences in Africa and Europe compared to other regions.}
    \label{fig:pca_results}
\end{figure}

>>>>

The Julia interface mainly defines the format of input data, so all in situ Julia data analysis methods can be plugged-in without additional E3SM compilation, thanks to Julia's Just-in-time (JIT) compilation.
As discussed previously, another key design of the Julia interface is to correctly convert the Fortran MPI communicator to Julia MPI communicator, which enables the algorithm developers to write their analysis routines using the same EAM MPI communicator. Due to the low-level MPI implementations, different MPI libraries (i.e., OpenMPI, MPICH, and MVAPICH) require specific converters; the following code snippet shows the converter implementations~\cite{gabriel2004open,panda2021mvapich,gropp1996user}.

@@

\subsection{PCA Results}
Over the one-year simulation run, we extracted the top three spatial principal components of the surface temperature variable using the TributaryPCA method; the principal components were updated in a streaming fashion at each simulation time step (6-hour time steps).
Figure \ref{fig:pca_results} shows spatial plots of these three principal components. 
Each component indicates a different spatial pattern of warming (red color) and cooling (blue color) extracted from the simulation data.
In particular, the first component indicates a temperature differential between equatorial and polar regions; because the mean was not removed, this component can be interpreted as the mean pattern across the time steps. 
The second component indicates a temperature differential between much of the Eastern hemisphere and the Americas and parts of Africa. 
The third component indicates a temperature differential between Africa and Europe and other regions.
It is plausible that the second and third components correspond to the regular diurnal cycle, which is likely one of the largest sources of regular variation in the one-year timespan considered here.
While these components were extracted from a fairly short simulation run, the results demonstrate the ability of TributaryPCA to extract important climate patterns while running in a distributed in situ environment, without saving any raw data for post-processing.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{figures/insitu_pca.png}
    \caption{First three spatial principal components obtained by in situ TributaryPCA method. Within each component, the white color indicates no change to surface temperature, while blue and red indicate temperatures deviating in different directions from zero. The first principal component appears to indicate the overall difference between the poles and the equator. The second principal component indicates differences between the Eastern hemisphere and Africa/the Americas, while the third component indicates differences in Africa and Europe compared to other regions.}
    \label{fig:pca_results}
\end{figure}
